---
title: "Statistical analysis of a multipartite ecological network"
author: "Sophie Donnet, Pierre Barbillon"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
bibliography: biblio.bib
link-citations: yes
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Econetwork}
  %\VignetteEncoding{UTF-8}
---


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


<!-- badges: start -->
<!-- badges: end -->

The goal of GREMLIN is to perform statistical analysis of multipartite networks through a block model approach. 

Multipartite networks  consist in the joint observation of several networks implying some common individuals. The individuals (or entities represented by nodes)  at stake are partitioned into groups defined by their nature.
In what follows, these groups will be referred to as  **functional groups**.


# Installation

You can install the released version of GREMLIN  from [GitHub](https://github.com/Demiperimetre/GREMLIN) with:

```{r packages, eval=TRUE}
#devtools::install_github("Demiperimetre/GREMLIN")
library(GREMLIN)
```

# Mathematical Background

###  A collection of networks

Assume that $Q$ functional groups of individuals are at stake; Let $n_q$ be  the number of individuals  in the $q$-th functional group.

A multipartite network is a collection of networks:  each network may be simple (relations inside  a   functional   group)   or bipartite  (relations between individuals of two  functional groups). We index the collection of networks by pairs of functional groups $(q,q')$. 

The set $E$ denotes  the  list of pairs of  functional groups  for   which we observe an interaction network.

For any pair $(q,q') \in E$, the interaction network is encoded in a matrix  $X^{qq'}$  :  $X^{qq'}_{ii'} \neq 0 $ if there is an edge from unit $i$ of functional group
$q$ to unit $i'$ of functional group $q'$, $0$ otherwise.
    - If $q \neq q'$,   $X^{qq'}$ is  said to be  an \emph{incidence matrix}.
    - $X^{qq}$ is  an \emph{adjacency matrix}:   it is symmetric if the relation inside the functional group $q$  is non-oriented, non-symmetric otherwise.

### A probabilistic latent variables model


Assume that,   each functional group $q$  is  divided into $K_q$  blocks (or equivalently clusters).  $\forall q $ and  $ \forall i $, let $Z^{q}_i$ be
the latent random variable  such that $Z^ q_i =k$  if individual $i$ of functional group $q$ belongs to cluster $k$.
The random variables $Z^{q}_i$'s are assumed to be independent and  such that: $\forall k, \forall q , \forall i $:

\begin{equation}\label{eq:mod2}
\mathbb{P}(Z^{q}_i=k) = \pi^{q}_k,
\end{equation}
with $\sum_{k=1}^{K_q}\pi^{q}_k=1$,  $\forall q$.



Conditionally on the clustering,  the entries of the matrices   $   \left(X^{qq'}_{ii'}\right)$ are assumed to be independent and distributed as follows:
$\forall (i,i')$,
\begin{equation}\label{eq:mod1}
X^{qq'}_{ii'} | \{Z^{q}_i=k, Z^{q'}_{i'}=k'  \} \sim_{ind} \mathcal{F}_{qq'}(\theta^{qq'}_{kk'})\,
\end{equation}
meaning that the probability of connection from $i$  of functional group $q$ to  $i'$ of functional group $q'$ only  depends on the clusters to which they belong to.\\

For any pair $(q,q')$, $\mathcal{F}_{qq'}(\cdot)$ is either: 

    - Bernoulli, resulting into binary interactions
    
    - Poisson for weighted networks of counts
    
    - Gaussian or Laplace for continuous weighted networks. 


The inference of the model consists in the selection of the number of clusters $(K_q)$ and the estimation of the parameters $(\theta_{qq'})$. The model selection is performed with the ICL, a penalized likelihood criterion. The parameters are estimated with a varitional version of the EM algorithm. The estimation procedure also provides a clustering of the entities at stake. 




# Example of a mutualistic ecological network

## The dataset
We apply our methodology to an ecological mutualistic multipartite network (see paper @Dattilo).

In this example, the functional groups are plants, floral visitors, ants and birds. We observe binary interactions between plants and floralvisitor,  plants and ants, and plants and seed dispersal birds.  


```{r loading dataset, eval=TRUE}
data(MPEcoNetwork, package = "GREMLIN")
```
These 3 networks are provided in 3 incidence matrices, the plants being in rows. *Note that the order of the individuals within the functional groups must be the same in all the matrices*.  


GREMLIN requires  the global network  be encoded in separate matrices for each network. $3$ incidence matrices. We then format the data to be able to use our R package GREMLIN i.e. we transform the matrices into an list containing    the matrix, its type ( `inc`: incidence, `adj`: adjacency symmetric, `dir_adj`: adjacency non symmetric, the functional group in row, the functional group in col). 


```{r transform dataset,  eval=TRUE}
PlantFlovis = defineNetwork(MPEconetwork$Inc_plant_flovis,"inc","plants","flovis")
PlantAnt = defineNetwork(MPEconetwork$Inc_plant_ant,"inc","plants","ants")
PlantBird = defineNetwork(MPEconetwork$Inc_plant_bird,"inc","plants","birds")
list_net <- list(PlantFlovis,PlantAnt,PlantBird)
names(PlantFlovis)
```

If one wants to keep a track of the names of the species, they should be used as rownames and colnames in the matrices. 

```{r example of dataset, eval=TRUE}
PlantFlovis$mat[1:2,1:2]

```

## Estimation 

The model selection and the estimation are performed with the function `multipartiteBM`.

```{r MBM, echo = FALSE, eval = TRUE}
if (!file.exists('resMBM_Mutualistic.Rda')) {
  RES_MBM = multipartiteBM(
    list_Net = list(PlantFlovis, PlantAnt, PlantBird),
    namesFG = c('plants','flovis','ants','birds'),
    v_distrib  = c('bernoulli','bernoulli','bernoulli'),
    initBM = TRUE,
    save = TRUE)
  save(RES_MBM,file="resMBM_Mutualistic.Rda")
} else {load("resMBM_Mutualistic.Rda")}
```

```{r MBM eval false, echo = TRUE, eval = FALSE}
RES_MBM = multipartiteBM(
    list_Net =list(PlantFlovis, PlantAnt, PlantBird),
    namesFG = c('plants','flovis','ants','birds'),
    v_distrib  = c('bernoulli','bernoulli','bernoulli'),
    initBM = TRUE,save = TRUE)
```



RES_MBM contains the estimated parameters of the models we run through during the search of the better number of blocks. 
If one sets `save = FALSE`  in the `multipartiteBM` function then we only save the best model. 

RES_MBM constains de dataset and the results.  
```{r MBM what}
names(RES_MBM)
``` 

The better model has the following numbers of blocks
```{r MBM v_K }
RES_MBM$fittedModel[[1]]$paramEstim$v_K
```

To see the parameters estimated for the better model we use the following command `RES_MBM$fittedModel[[1]]$paramEstim$***`
```{r MBM param }
RES_MBM$fittedModel[[1]]$paramEstim$list_pi$plants
RES_MBM$fittedModel[[1]]$paramEstim$list_theta$plantsflovis
```

The clustering  supplied by the better model  are in `RES_MBM$fittedModel[[1]]$paramEstim$Z$***`.

```{r MBM Z }
table(RES_MBM$fittedModel[[1]]$paramEstim$Z$plants)
table(RES_MBM$fittedModel[[1]]$paramEstim$Z$ants)      
``` 


# References
