---
title: "Illustration of the GREMLIN package on a simulated dataset"
author: "Sophie Donnet, Pierre Barbillon"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
bibliography: biblio.bib
link-citations: yes
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Econetwork}
  %\VignetteEncoding{UTF-8}
---


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


<!-- badges: start -->
<!-- badges: end -->

The goal of GREMLIN is to perform statistical analysis of multipartite networks through a block model approach. 

Multipartite networks  consist in the joint observation of several networks implying some common individuals. The individuals (or entities represented by nodes)  at stake are partitioned into groups defined by their nature.
In what follows, these groups will be referred to as  **functional groups**.


# Installation

You can install the released version of GREMLIN  from [GitHub](https://github.com/Demiperimetre/GREMLIN) with:

```{r packages, eval=TRUE}
#devtools::install_github("Demiperimetre/GREMLIN")
library(GREMLIN)
```

# Mathematical Background

###  A collection of networks

Assume that $Q$ functional groups of individuals are at stake; Let $n_q$ be  the number of individuals  in the $q$-th functional group.

A multipartite network is a collection of networks:  each network may be simple (relations inside  a   functional   group)   or bipartite  (relations between individuals of two  functional groups). We index the collection of networks by pairs of functional groups $(q,q')$. 

The set $E$ denotes  the  list of pairs of  functional groups  for   which we observe an interaction network.

For any pair $(q,q') \in E$, the interaction network is encoded in a matrix  $X^{qq'}$  :  $X^{qq'}_{ii'} \neq 0 $ if there is an edge from unit $i$ of functional group
$q$ to unit $i'$ of functional group $q'$, $0$ otherwise.
    - If $q \neq q'$,   $X^{qq'}$ is  said to be  an \emph{incidence matrix}.
    - $X^{qq}$ is  an \emph{adjacency matrix}:   it is symmetric if the relation inside the functional group $q$  is non-oriented, non-symmetric otherwise.

### A probabilistic latent variables model


Assume that,   each functional group $q$  is  divided into $K_q$  blocks (or equivalently clusters).  $\forall q $ and  $ \forall i $, let $Z^{q}_i$ be
the latent random variable  such that $Z^ q_i =k$  if individual $i$ of functional group $q$ belongs to cluster $k$.
The random variables $Z^{q}_i$'s are assumed to be independent and  such that: $\forall k, \forall q , \forall i $:

\begin{equation}\label{eq:mod2}
\mathbb{P}(Z^{q}_i=k) = \pi^{q}_k,
\end{equation}
with $\sum_{k=1}^{K_q}\pi^{q}_k=1$,  $\forall q$.



Conditionally on the clustering,  the entries of the matrices   $   \left(X^{qq'}_{ii'}\right)$ are assumed to be independent and distributed as follows:
$\forall (i,i')$,
\begin{equation}\label{eq:mod1}
X^{qq'}_{ii'} | \{Z^{q}_i=k, Z^{q'}_{i'}=k'  \} \sim_{ind} \mathcal{F}_{qq'}(\theta^{qq'}_{kk'})\,
\end{equation}
meaning that the probability of connection from $i$  of functional group $q$ to  $i'$ of functional group $q'$ only  depends on the clusters to which they belong to.\\

For any pair $(q,q')$, $\mathcal{F}_{qq'}(\cdot)$ is either: 

    - Bernoulli, resulting into binary interactions
    
    - Poisson for weighted networks of counts
    
    - Gaussian or Laplace for continuous weighted networks. 


The inference of the model consists in the selection of the number of clusters $(K_q)$ and the estimation of the parameters $(\theta_{qq'})$. The model selection is performed with the ICL, a penalized likelihood criterion. The parameters are estimated with a varitional version of the EM algorithm. The estimation procedure also provides a clustering of the entities at stake. 




# Simulation of a complex multipartite

We use the function  `rMBM` provided in the package to simulate a  multipartite network involving $3$ functional groups ('A','B', and 'C') of respective sizes $n_A = 100$, $n_B = 50$ and $n_C = 60$. 

A,B abd C are  devided respectively into $4$, $3$ and $2$ blocks. The sizes of the blocks are generated randomly. 

```{r param FG}
set.seed(302718)
n_FG <- 3 #number of functional groups
v_NQ = c(100,50,60)
v_K <- c(4,3,2) #number of clusters in each functional group
list_pi = lapply(1:n_FG,function(q){v = rgamma(v_K[q],1,1); return(v/sum(v))})
namesFG <- c('A','B','C')
``` 

We assume that we observe $4$ interactions matrices
    - A-B : continuous weighted interactions
    - B-C : binary interactions
    - A-A : counting directed interactions
    - C-C : binary non-directed interactions

```{r param net}
E  <-  rbind(c(1,2),c(2,3),c(1,1),c(3,3))
typeInter <- c( "inc",'inc',"diradj", "adj")
v_distrib <- c('gaussian','bernoulli','poisson','bernoulli')
list_theta <- list()
list_theta[[1]] <- list()
list_theta[[1]]$mean  <- matrix(round(rnorm(v_K[E[1,1]] * v_K[E[1,2]],7.5,4 ),1),nrow = v_K[E[1,1]], ncol = v_K[E[1,2]] )
list_theta[[1]]$var  = matrix(round(rgamma(v_K[E[1,1]] * v_K[E[1,2]],7.5,4 ),1),nrow = v_K[E[1,1]], ncol = v_K[E[1,2]] )
for (q in 2:4) {
  list_theta[[q]] <- switch(v_distrib[q], 
      bernoulli = matrix(rbeta(v_K[E[q,1]] * v_K[E[q,2]],2,2 ),nrow = v_K[E[q,1]], ncol = v_K[E[q,2]]),
      poisson = matrix(rgamma(v_K[E[q,1]] * v_K[E[q,2]],6,2 ),nrow = v_K[E[q,1]], ncol = v_K[E[q,2]])
  )
}
list_theta[[4]] =0.5 * (list_theta[[4]] + t(list_theta[[4]]))

``` 

We are now ready to simulate the data

```{r simul, eval = TRUE}
dataSim <- rMBM(v_NQ ,E , typeInter, v_distrib, list_pi, 
                list_theta, seed = NULL, 
                namesFG =namesFG,keepClassif  = TRUE)
list_Net <- dataSim$list_Net
length(list_Net)

```



## Estimation 

The model selection and the estimation are performed with the function `multipartiteBM`.


```{r MBM, echo = FALSE, eval = TRUE}
if (!file.exists('resMBM_Simu.Rda')) {
  res_MBMsimu <- multipartiteBM(list_Net, 
                      v_distrib = v_distrib, 
                      v_Kmin = 1,
                      v_Kmax = 10,
                      v_Kinit = NULL,
                      verbose = TRUE, 
                      save=FALSE, initBM = TRUE)
  save(RES_MBM,file="resMBM_Simu.Rda")
} else {load("resMBM_Simu.Rda")}
```

```{r MBM eval false, echo = TRUE, eval = FALSE}
  res_MBMsimu <- multipartiteBM(list_Net, 
                      v_distrib = v_distrib, 
                      v_Kmin = 1,
                      v_Kmax = 10,
                      v_Kinit = NULL,
                      verbose = TRUE, 
                      save=FALSE, initBM = TRUE)
```

We can now get the estimated parameters and compare them to the true ones. 

```{r estim param}


res_MBMsimu$fittedModel[[1]]$paramEstim$list_theta$AB$mean
list_theta[[1]]$mean

```


# References
